Fundamentals of Computing and Data Display

### 2019-12-07 Paper/paper-draft-Padgett-Thiam.Rmd
update paper code and add intro

---
title: "Fundamentals of Computing and Data Display"
subtitle: "Term paper template"
author: "Author"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    df_print: kable
references:
- id: Wickham2014
  title: Tidy Data
  author:
  - family: Wickham
    given: Hadley
  container-title: Journal of Statistical Software
  volume: 59
  issue: 10
  page: 1-23
  type: article-journal
  issued:
    year: 2014
- id: Baumer2017
  title: Modern Data Science with R
  author:
  - family: Baumer
    given: Benjamin S.
  - family: Kaplan
    given: Daniel T.
  - family: Horton
    given: Nicholas J.
  type: book
  publisher: Chapman \& Hall/CRC Press.
  issued:
    year: 2017
---

TO DO:
- write intro, results, and discussion sections
- save the following datasets to github and write code in designated chunks to bring them in:
    props_tab1
    props_tab2
    gc_locations
    interest_over_time_all
    pre_sent_pos
    post_sent_pos
- create presentation

@@

This section outlines the research idea. We can also cite related work here [@Wickham2014; @Baumer2017].

Note that compiled term paper (the PDF) is supposed to be more text-centered than the RMarkdown documents we used in class, i.e. the text sections are more detailed and big or redundant code chunks can be hidden.

@@

```{r, eval=FALSE}

@@

Next, we seperate the tweets by candidates in order to do the analysis at the candidate level. It's important to note that one tweets can be addressed to multiple candidates. In that case, the tweet will be found in the each of those candidate dataset.

```{r, eval=FALSE}
tweets$BS <-(str_detect(tweets$text, regex("#Sanders|@BernieSanders", ignore_case = TRUE)))
tweets$KH <-(str_detect(tweets$text, regex("#harris |@KamalaHarris",  ignore_case = TRUE)))
tweets$JB <-(str_detect(tweets$text, regex("#biden |@JoeBiden", ignore_case = TRUE)))
tweets$EW <-(str_detect(tweets$text, regex("#warren |@ewarren", ignore_case = TRUE)))
tweets$PB <-(str_detect(tweets$text, regex("#buttigieg|@PeteButtigieg", ignore_case = TRUE)))
tweets$TS <-(str_detect(tweets$text, regex("#steyer | @TomSteyer", ignore_case = TRUE)))
tweets$AY <-(str_detect(tweets$text, regex("#yang| @AndrewYang", ignore_case = TRUE)))
tweets$BC <-(str_detect(tweets$text, regex("#booker | @BookerCory", ignore_case = TRUE)))
tweets$AK <-(str_detect(tweets$text, regex("#klobuchar | @amyklobuchar", ignore_case = TRUE)))
tweets$TG <-(str_detect(tweets$text, regex("#gabbard |@TulsiGabbard", ignore_case = TRUE)))

@@

```{r, eval=FALSE}
df1$BS <-(str_detect(df1$text, regex("#Sanders|@BernieSanders", ignore_case = TRUE)))
df1$KH <-(str_detect(df1$text, regex("#harris |@KamalaHarris",  ignore_case = TRUE)))
df1$JB <-(str_detect(df1$text, regex("#biden |@JoeBiden", ignore_case = TRUE)))
df1$EW <-(str_detect(df1$text, regex("#warren |@ewarren", ignore_case = TRUE)))
df1$PB <-(str_detect(df1$text, regex("#buttigieg|@PeteButtigieg", ignore_case = TRUE)))
df1$TS <-(str_detect(df1$text, regex("#steyer | @TomSteyer", ignore_case = TRUE)))
df1$AY <-(str_detect(df1$text, regex("#yang| @AndrewYang", ignore_case = TRUE)))
df1$BC <-(str_detect(df1$text, regex("#booker | @BookerCory", ignore_case = TRUE)))
df1$AK <-(str_detect(df1$text, regex("#klobuchar | @amyklobuchar", ignore_case = TRUE)))
df1$TG <-(str_detect(df1$text, regex("#gabbard |@TulsiGabbard", ignore_case = TRUE)))

@@

```{r, eval=FALSE}
df2$BS <-(str_detect(df2$text, regex("#Sanders|@BernieSanders", ignore_case = TRUE)))
df2$KH <-(str_detect(df2$text, regex("#harris |@KamalaHarris",  ignore_case = TRUE)))
df2$JB <-(str_detect(df2$text, regex("#biden |@JoeBiden", ignore_case = TRUE)))
df2$EW <-(str_detect(df2$text, regex("#warren |@ewarren", ignore_case = TRUE)))
df2$PB <-(str_detect(df2$text, regex("#buttigieg|@PeteButtigieg", ignore_case = TRUE)))
df2$TS <-(str_detect(df2$text, regex("#steyer | @TomSteyer", ignore_case = TRUE)))
df2$AY <-(str_detect(df2$text, regex("#yang| @AndrewYang", ignore_case = TRUE)))
df2$BC <-(str_detect(df2$text, regex("#booker | @BookerCory", ignore_case = TRUE)))
df2$AK <-(str_detect(df2$text, regex("#klobuchar | @amyklobuchar", ignore_case = TRUE)))
df2$TG <-(str_detect(df2$text, regex("#gabbard |@TulsiGabbard", ignore_case = TRUE)))

@@

All the tweets received before 9 p.m. on November 20th are accounted for in the pre-debate dataset and all tweets from 11 p.m. on November 20th to the next day are in the post-debate dataset. Note that tweets during the debate( 9 - 11 p.m are ignored)
```{r, eval=FALSE}
pre_debate <- tweets %>%
              filter(date(date) == "2019-11-20" & hour(date) < 21 )

@@

First we need to prep the tweets by removing all non-words such as emojis and use the sentiment analysis package for analysis. Althought the sentiment analysis packae uses 5 dictionnairies, we will only look at the results from the GI dictionnary. For the pre-debate tweets sentimeent we used all the tweets from the dataset; however for the post-debate analysis, we sampled from the dataset due to volume and processing error.   
```{r, eval=FALSE}
bs <- pre_debate %>%
      filter (BS == "TRUE")

@@

bs_post <- post_debate %>%
      filter (BS == "TRUE")

@@

```{r, eval=FALSE}
kh <- pre_debate %>%
      filter (KH == "TRUE")

@@

```{r, eval=FALSE}
jb <- pre_debate %>%
      filter (JB == "TRUE")

@@

```{r, eval=FALSE}
ew <- pre_debate %>%
      filter (EW == "TRUE")

@@

```{r, eval=FALSE}
pb <- pre_debate %>%
      filter (PB == "TRUE")

@@

```{r, eval=FALSE}
pre_sent_pos <- data.frame("Candidate" = c("Biden", "Sanders", "Warren", "Harris", "Buttigieg"), 
                       "Positive" =c((sum(sentiments_jb$PositivityGI))/27477,
                                      (sum(sentiments_bs$PositivityGI))/31603,                                                                  (sum(sentiments_ew$PositivityGI))/18628,
                                     (sum(sentiments_kh$PositivityGI))/16362,
                                      (sum(sentiments_pb$PositivityGI))/27242))
```

```{r, eval=FALSE}
post_sent_pos <- data.frame("Candidate" = c("Biden", "Sanders", "Warren", "Harris", "Buttigieg"), 
                       "Positive" =c((sum(sentiments_jb_post$PositivityGI))/27477,
                                      (sum(sentiments_bs_post$PositivityGI))/31603,                                                             (sum(sentiments_ew_post$PositivityGI))/18628,
                                     (sum(sentiments_kh_post$PositivityGI))/16362,
                                      (sum(sentiments_pb_post$PositivityGI))/27242))
```

@@

```{r, eval=FALSE}
register_google(key = "XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX")
res1 <- gtrends(c("Joe Biden", "Bernie Sanders", "Elizabeth Warren", "Kamala Harris", "Pete Buttigieg"), geo = "US", 
                time = "2019-11-18 2019-11-22", low_search_volume = T)

@@

```{r}
github_link <- "https://github.com/znpadgett/surv727_padgett_thiam/raw/master/Data/Project%20polling%20data.xlsx"
temp_file <- tempfile(fileext = ".xlsx")
req <- GET(github_link, 
           write_disk(path = temp_file))
polling_data <- readxl::read_excel(temp_file)
```

@@

```{r}
#get map
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
us_map <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
ggmap(us_map)

#clean data for mapping
trends_map <-
  gc_locations %>%
  group_by(keyword, location) %>%
  mutate(total=sum(hits))

@@

```{r}
interest_over_time_all %>%
  group_by(keyword, Pre_post) %>%
  summarise(total_hits=sum(hits)) %>%
  ggplot() +
  geom_col(mapping = aes(x=reorder(keyword, -total_hits), y=total_hits, fill=Pre_post), color="black")
```

@@

For the polling data, we looked at the pre- and post-debate polling numbers for each candidate.
```{r}
polling_data %>%
  group_by(Candidate, Pre_post) %>%
  ggplot() +
    geom_col(mapping = aes(x=reorder(Candidate, -Percentage), y=Percentage, fill=Pre_post), color="black",   position="dodge")
```

@@

```{r}
#clean data
gtrends <-
  interest_over_time_all %>%
  mutate(Candidate=keyword, Data="GTrends") %>%
  group_by(Candidate, Pre_post, Data) %>%
  summarise(Percentage=mean(hits))

@@

Then, we created a shiny app, which can be viewed by accessing the following site: ________________.
```{r}
# Define UI
ui <- fluidPage(

>>>>

---
title: "Fundamentals of Computing and Data Display"
subtitle: "Term paper template"
author: "Author"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
    df_print: kable
references:
- id: GayoAvello2013
  title: A Meta-Analysis of State-of-the-Art Electoral Prediction From Twitter Data
  author:
  - family: Gayo-Avello
    given: D.
  container-title: Organization Studies
  volume: 31
  issue: 6
  page: 211–248
  type: article-journal
  issued:
    year: 2013
- id: Beauchamp2017
  title: Predicting and Interpolating State-Level Polls Using Twitter Textual Data
  author:
  - family: Beauchamp
    given: N.
  container-title: American Journal of Political Science 
  volume: 61
  issue: 2
  type: article-journal
  issued:
    year: 2017
- id: Gaurav2013
  title: Leveraging Candidate Popularity on Twitter to Predict Election Outcome
  author:
  - family: Gaurav
    given: M.
  - family: Srivastava
    given: A.
  - family: Kumar
    given: A.
  - family: Miller
    given: S.
  container-title: Proceedings of the 7th Workshop on Social Network Mining and Analysis
  type: article-journal
  issued:
    year: 2013
- id: Kassraie2017
  title: Election Vote Share Prediction using a Sentiment-based Fusion of Twitter Data with Google Trends and Online Polls
  author:
  - family: Kassraie
    given: P.
  - family: Modirshanechi
    given: A.
  - family: Aghajan
    given: H.
  container-title: Proceedings of the 6th International Conference on Data Science, Technology and Applications
  type: article-journal
  issued:
    year: 2017
- id: Salunkhe2017
  title: Twitter Based Election Prediction and Analysis
  author:
  - family: Salunkhe
    given: P.
  - family: Deshmukh
    given: S.
  container-title: International Research Journal of Engineering and Technology
  volume: 4
  issue: 10
  type: article-journal
  issued:
    year: 2017
---

TO DO:
- write results, and discussion sections
- create presentation

@@

Twitter and Google Trends are becoming increasingly popular tools for social science researchers. They represent an easily accessible source of large amounts of data, which has become advantageous as survey response rates decline and costs rise. Researchers interested in predicting election results have begun looking to Twitter data to replace or supplement traditional election polls, with mixed results [@GayoAvello2013].

Recently, there have been studies using sentiment analysis of Twitter data to predict election outcomes in India [@Salunkhe2017], to predict state-level polling results in the U.S. [@Beauchamp2017], and to predict the winners of three presidential elections in Latin America [@Gaurav2013]. Beauchamp (2017) found that Twitter data may be useful in making state-level campaign strategy decisions. Additionally, a recent study used Google Trends and Twitter data to predict the 2016 U.S. election outcomes with only 1% error [@Kassraie2017].

We are interested in whether Twitter data and Google trends data could be used to supplement polling results, by providing real-time information to candidates while they wait for polling data to come in. For example, candidates may be interested in understanding how public opinion has shifted immediately after a debate in order to run a more agile campaign. This project is an exploratory analysis of Twitter and Google Trends data to see if pre- and post-debate polling data during the 2020 U.S. Democratic primary election aligns with these real-time sources.

@@

```{r, eval=FALSE, tidy=TRUE}

@@

Next, we seperate the tweets by candidates in order to do the analysis at the candidate level. It's important to note that one tweets can be addressed to multiple candidates. In that case, the tweet will be found in the each of those candidate dataset. The code below exemplifies this step of the cleaning process.

```{r, eval=FALSE, tidy=TRUE}
tweets$BS <-(str_detect(tweets$text, regex("#Sanders|@BernieSanders", ignore_case = TRUE)))
tweets$KH <-(str_detect(tweets$text, regex("#harris |@KamalaHarris",  ignore_case = TRUE)))
tweets$JB <-(str_detect(tweets$text, regex("#biden |@JoeBiden", ignore_case = TRUE)))
tweets$EW <-(str_detect(tweets$text, regex("#warren |@ewarren", ignore_case = TRUE)))
tweets$PB <-(str_detect(tweets$text, regex("#buttigieg|@PeteButtigieg", ignore_case = TRUE)))
tweets$TS <-(str_detect(tweets$text, regex("#steyer | @TomSteyer", ignore_case = TRUE)))
tweets$AY <-(str_detect(tweets$text, regex("#yang| @AndrewYang", ignore_case = TRUE)))
tweets$BC <-(str_detect(tweets$text, regex("#booker | @BookerCory", ignore_case = TRUE)))
tweets$AK <-(str_detect(tweets$text, regex("#klobuchar | @amyklobuchar", ignore_case = TRUE)))
tweets$TG <-(str_detect(tweets$text, regex("#gabbard |@TulsiGabbard", ignore_case = TRUE)))

@@

```{r, eval=FALSE, include=FALSE}
df1$BS <-(str_detect(df1$text, regex("#Sanders|@BernieSanders", ignore_case = TRUE)))
df1$KH <-(str_detect(df1$text, regex("#harris |@KamalaHarris",  ignore_case = TRUE)))
df1$JB <-(str_detect(df1$text, regex("#biden |@JoeBiden", ignore_case = TRUE)))
df1$EW <-(str_detect(df1$text, regex("#warren |@ewarren", ignore_case = TRUE)))
df1$PB <-(str_detect(df1$text, regex("#buttigieg|@PeteButtigieg", ignore_case = TRUE)))
df1$TS <-(str_detect(df1$text, regex("#steyer | @TomSteyer", ignore_case = TRUE)))
df1$AY <-(str_detect(df1$text, regex("#yang| @AndrewYang", ignore_case = TRUE)))
df1$BC <-(str_detect(df1$text, regex("#booker | @BookerCory", ignore_case = TRUE)))
df1$AK <-(str_detect(df1$text, regex("#klobuchar | @amyklobuchar", ignore_case = TRUE)))
df1$TG <-(str_detect(df1$text, regex("#gabbard |@TulsiGabbard", ignore_case = TRUE)))

@@

```{r, eval=FALSE, include=FALSE}
df2$BS <-(str_detect(df2$text, regex("#Sanders|@BernieSanders", ignore_case = TRUE)))
df2$KH <-(str_detect(df2$text, regex("#harris |@KamalaHarris",  ignore_case = TRUE)))
df2$JB <-(str_detect(df2$text, regex("#biden |@JoeBiden", ignore_case = TRUE)))
df2$EW <-(str_detect(df2$text, regex("#warren |@ewarren", ignore_case = TRUE)))
df2$PB <-(str_detect(df2$text, regex("#buttigieg|@PeteButtigieg", ignore_case = TRUE)))
df2$TS <-(str_detect(df2$text, regex("#steyer | @TomSteyer", ignore_case = TRUE)))
df2$AY <-(str_detect(df2$text, regex("#yang| @AndrewYang", ignore_case = TRUE)))
df2$BC <-(str_detect(df2$text, regex("#booker | @BookerCory", ignore_case = TRUE)))
df2$AK <-(str_detect(df2$text, regex("#klobuchar | @amyklobuchar", ignore_case = TRUE)))
df2$TG <-(str_detect(df2$text, regex("#gabbard |@TulsiGabbard", ignore_case = TRUE)))

@@

All the tweets received before 9 p.m. on November 20th are accounted for in the pre-debate dataset and all tweets from 11 p.m. on November 20th to the next day are in the post-debate dataset. Note that tweets during the debate( 9 - 11 p.m are ignored)
```{r, eval=FALSE, tidy=TRUE}
pre_debate <- tweets %>%
              filter(date(date) == "2019-11-20" & hour(date) < 21 )

@@

First we need to prep the tweets by removing all non-words such as emojis and use the sentiment analysis package for analysis. Althought the sentiment analysis packae uses 5 dictionnairies, we will only look at the results from the GI dictionnary. For the pre-debate tweets sentimeent we used all the tweets from the dataset; however for the post-debate analysis, we sampled from the dataset due to volume and processing error. The code below exemplifies this process.   
```{r, eval=FALSE}
bs <- pre_debate %>%
      filter (BS == "TRUE")

@@

```{r, eval=FALSE, include=FALSE}
bs_post <- post_debate %>%
      filter (BS == "TRUE")

@@

```{r, eval=FALSE, include=FALSE}
kh <- pre_debate %>%
      filter (KH == "TRUE")

@@

```{r, eval=FALSE, include=FALSE}
jb <- pre_debate %>%
      filter (JB == "TRUE")

@@

```{r, eval=FALSE, include=FALSE}
ew <- pre_debate %>%
      filter (EW == "TRUE")

@@

```{r, eval=FALSE, include=FALSE}
pb <- pre_debate %>%
      filter (PB == "TRUE")

@@

```{r, eval=FALSE, tidy=TRUE}
pre_sent_pos <- data.frame("Candidate" = c("Biden", "Sanders", "Warren", "Harris", "Buttigieg"), 
                       "Positive" =c((sum(sentiments_jb$PositivityGI))/27477,
                                      (sum(sentiments_bs$PositivityGI))/31603,                                                                  (sum(sentiments_ew$PositivityGI))/18628,
                                     (sum(sentiments_kh$PositivityGI))/16362,
                                      (sum(sentiments_pb$PositivityGI))/27242))
```

```{r, eval=FALSE, include=FALSE}
post_sent_pos <- data.frame("Candidate" = c("Biden", "Sanders", "Warren", "Harris", "Buttigieg"), 
                       "Positive" =c((sum(sentiments_jb_post$PositivityGI))/27477,
                                      (sum(sentiments_bs_post$PositivityGI))/31603,                                                             (sum(sentiments_ew_post$PositivityGI))/18628,
                                     (sum(sentiments_kh_post$PositivityGI))/16362,
                                      (sum(sentiments_pb_post$PositivityGI))/27242))
```

@@

```{r, eval=FALSE, tidy=TRUE}
register_google(key = "XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX")
res1 <- gtrends(c("Joe Biden", "Bernie Sanders", "Elizabeth Warren", "Kamala Harris", "Pete Buttigieg"), geo = "US", 
                time = "2019-11-18 2019-11-22", low_search_volume = T)

@@

```{r, tidy=TRUE}
github_link <- "https://github.com/znpadgett/surv727_padgett_thiam/raw/master/Data/Project%20polling%20data.xlsx"
temp_file <- tempfile(fileext = ".xlsx")
req <- GET(github_link, 
           write_disk(path = temp_file))
polling_data <- readxl::read_excel(temp_file)
```

@@

```{r, message=FALSE, warning=FALSE, results="hide", fig.show="hide"}
#get map
us <- c(left = -125, bottom = 25.75, right = -67, top = 49)
us_map <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
ggmap(us_map) 
```

```{r, message=FALSE, warning=FALSE}
#clean data for mapping
trends_map <-
  gc_locations %>%
  group_by(keyword, location) %>%
  mutate(total=sum(hits))

@@

```{r, tidy=TRUE}
interest_over_time_all %>%
  group_by(keyword, Pre_post) %>%
  summarise(total_hits=sum(hits)) %>%
  ggplot() +
  geom_col(mapping = aes(x=reorder(keyword, -total_hits), y=total_hits, fill=Pre_post), color="black")
```

@@

For the polling data, we looked at the pre- and post-debate polling numbers for each candidate.
```{r, tidy=TRUE}
polling_data %>%
  group_by(Candidate, Pre_post) %>%
  ggplot() +
    geom_col(mapping = aes(x=reorder(Candidate, -Percentage), y=Percentage, fill=Pre_post), color="black",   position="dodge")
```

@@

```{r, tidy=TRUE, warning=FALSE, message=FALSE}
#clean data
gtrends <-
  interest_over_time_all %>%
  mutate(Candidate=keyword, Data="GTrends") %>%
  group_by(Candidate, Pre_post, Data) %>%
  summarise(Percentage=mean(hits))

@@

Then, we created a shiny app, which can be viewed by accessing the following site: ________________.
```{r, eval=FALSE, tidy=TRUE}
# Define UI
ui <- fluidPage(
